Thomas et al. 2025

### 2025-01-10 AT_2025.qmd
New formatting .R file

## Read the data to a df
AT_2025_data <- readr::read_csv("DataTables/AT_ChemicalCommunications_2025.csv")

## Pivot the data to be longer
AT_2025_Long <- AT_2025_data |>
	pivot_longer(cols = -gene_name,
				names_to = c(".value", "Sample"),
				names_pattern = "^(logFC|pvalue|hit_annotation)_(.+)$")

@@

ID <- c()
for(i in 0:(nrow(AT_2025_Long)/2)){
	ID[i] = i
}

### The ggplot customizations that make the ranked order plots
AT_2025_ranked_PE <- AT_2025_Long |>
    arrange(logFC) |>
	filter(Sample == "PE") |>
	mutate(ID = ID)

AT_2025_ranked_PA <- AT_2025_Long |>
    arrange(logFC) |>
	filter(Sample == "PA") |>
	mutate(ID = ID)

AT_2025_ranked <- rbind(AT_2025_ranked_PE, AT_2025_ranked_PA)

?merge.data.frame()

RankedOrderPlots <- AT_2025_ranked |>
    ggplot(aes(
        x = ID,
        y = logFC,
        color = factor(hit_annotation),
        fill = factor(hit_annotation),
        shape = factor(hit_annotation),
        alpha = factor(hit_annotation),
        size = factor(hit_annotation)
    )) +
    geom_point(aes(text = paste0("Gene name: ", gene_name, "\n",
                                 "LogFC: ", logFC, "\n",
                                 "p-value: ", pvalue, "\n"))) +
    geom_hline(yintercept = 0, linetype = 2) +
    geom_vline(xintercept = 0, linetype = 2) +
    scale_shape_manual(values = c("enriched hit" = 21,
                                  "enriched candidate" = 24,
                                  "no hit" = 22), 
                       name = "Trend") +
    scale_color_manual(values = c("enriched hit" = "black",
                                  "enriched candidate" = "black",
                                  "no hit" = "black"), 
                       name = "Trend") +
    scale_fill_manual(values = c("enriched hit" = "orange",
                                 "enriched candidate" = "purple",
                                 "no hit" = "black"), 
                      name = "Trend") +
    scale_alpha_manual(values = c("enriched hit" = 0.8,
                                  "enriched candidate" = 0.5,
                                  "no hit" = 0.25), 
                       name = "Trend") +
    scale_size_manual(values = c("enriched hit" = 4,
                                 "enriched candidate" = 2,
                                 "no hit" = 0.75), 
                      name = "Trend") +
    facet_wrap(~Sample) +
	ylab("Log2 fold-change") +
	xlab("Ranked-order gene") +
    theme_minimal() +
    theme(panel.border = element_rect(color = "#cfcfcf", fill = NA)) +
    theme(legend.title = element_text(size = 12, face = "bold")) +
    theme(axis.title = element_text(size = 12, face = "bold")) +
    theme(legend.text = element_text(size = 12))

m <- list(l=50, r = 50, b= 100, pad = 4)

RankedOrderPlotly <- ggplotly(RankedOrderPlots,
						  tooltip = c("text", dynamicTicks = TRUE),
						  hovermode = "closest",
						  autosize = T,
						  margin = m) |>
				layout(showlegend = FALSE)

@@

:::{.border}
```{r, echo=FALSE}
#| fig-cap: "Ranked-order plots depict the relative enrichment of each protein versus control (y-axis) from lowest to highest. The left panel depicts the results of a pulldown using trifunctionalized phosphatidic acid; the right panel depicts the results of the pulldown using trifunctionalized phosphatidylethanolamine. Black proteins are unenriched or depleted in the presence of probe, Purple proteins are enriched to the probe but fall below a FDR or p-value threshold of significance, and Orange proteins surpass the threshold of significance (need details of threshold). "

RankedOrderPlotly
```
:::

### The ggplot customizations that make the volcano plots
VolcanoPlots <- AT_2025_Long |>
	ggplot(aes(x = logFC,
			   y = -log10(pvalue),
			   color = factor(hit_annotation), 
			   fill = factor(hit_annotation), 
			   shape = factor(hit_annotation),
			   alpha = factor(hit_annotation),
			   size = factor(hit_annotation))) +
	geom_point(aes(text = paste0("Gene name: ", gene_name, "\n",
								 "LogFC: ", logFC, "\n",
								 "p-value: ", pvalue, "\n"))) +
	geom_hline(yintercept = 0, linetype = 2) +
	geom_vline(xintercept = 0, linetype = 2) +
	scale_x_continuous(limits=c(-1.1 * max(abs(AT_2025_Long$logFC)), 1.1* max(abs(AT_2025_Long$logFC)))) +
    scale_y_continuous(limits=c(0, max(-log10(AT_2025_Long$pvalue)))) +
	scale_shape_manual(values = c("enriched hit" = 21,
								  "enriched candidate" = 24,
								  "no hit" = 22), 
					   name = "Trend") +
	scale_color_manual(values = c("enriched hit" = "black",
								 "enriched candidate" = "black",
								 "no hit" = "black"), 
								 name = "Trend") +
	scale_fill_manual(values = c("enriched hit" = "orange",
								  "enriched candidate" = "purple",
								  "no hit" = "black"), 
					   name = "Trend") +
	scale_alpha_manual(values = c("enriched hit" = 0.8,
								  "enriched candidate" = 0.5,
								  "no hit" = 0.25), 
					   name = "Trend") +
	scale_size_manual(values = c("enriched hit" = 4,
								  "enriched candidate" = 2,
								  "no hit" = 0.75), 
					   name = "Trend") +
	facet_wrap(~Sample) +
	xlab("Log2 fold-change") +
	ylab("-log10(p-value)") +
	theme_minimal() +
	theme(panel.border = element_rect(color = "#cfcfcf", fill = NA))+
	theme(strip.text = element_text(size = 12, face = "bold")) +
	theme(legend.title = element_text(size = 12, face = "bold")) +
	theme(axis.title = element_text(size = 12, face = "bold")) +
	theme(legend.text = element_text(size = 12))

m <- list(l=50, r = 50, b= 100, pad = 10)

VolcanoPlotly <- ggplotly(VolcanoPlots,
						  tooltip = c("text", dynamicTicks = TRUE),
						  hovermode = "closest",
						  autosize = T,
						  margin = m) |>
				layout(showlegend = FALSE)

VolcanoPlotly

@@

VolcanoPlotly
```
:::

>>>>

## Set up the source for the standardized ggplot styles
source("ggplot_styles.R")
```

# Data wrangling to prepare Alix's data to the standardized inputs for my plotting functions.
# Read the data to a df
AT_2025_data <- readr::read_csv("DataTables/AT_ChemicalCommunications_2025.csv")

# Pivot the data to be longer
AT_2025_Long <- AT_2025_data |>
	pivot_longer(cols = -gene_name,
				names_to = c(".value", "Sample"),
				names_pattern = "^(logFC|pvalue|hit_annotation)_(.+)$")

@@

# Building an ID column to maintain the order on the x-axis
ID <- c()
for(i in 0:(nrow(AT_2025_Long)/2)){
	ID[i] = i
}

# Applies the ID column to the PE results
AT_2025_ranked_PE <- AT_2025_Long |>
    arrange(logFC) |>
	filter(Sample == "PE") |>
	mutate(ID = ID)

# Applies the ID column to the PA results
AT_2025_ranked_PA <- AT_2025_Long |>
    arrange(logFC) |>
	filter(Sample == "PA") |>
	mutate(ID = ID)

# Assembles the dataset to plot
AT_2025_ranked <- rbind(AT_2025_ranked_PE, AT_2025_ranked_PA)

@@

:::{.border}
```{r, echo=FALSE, warning=FALSE}
#| fig-cap: "Ranked-order plots depict the relative enrichment of each protein versus control (y-axis) from lowest to highest. The left panel depicts the results of a pulldown using trifunctionalized phosphatidic acid; the right panel depicts the results of the pulldown using trifunctionalized phosphatidylethanolamine. Black proteins are unenriched or depleted in the presence of probe, Purple proteins are enriched to the probe but fall below a FDR or p-value threshold of significance, and Orange proteins surpass the threshold of significance (need details of threshold). "

RankedOrderPlotStandard(AT_2025_ranked)
```
:::

@@

VolcanoPlotStandardized(AT_2025_Long)
```
:::
